Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-02-08 text/src/article.tex
Two revisions of first background section

High throughput sequencing can be reads short regions of a genome sequence to
produce hundreds of thousands to millions of lines of nucleotide sequence. The
current status of parallel nucleotide sequencing is limited to producing reads
less than $>$1000 nucleotides long. Therefore to obtain a contiguous complete
genome sequence these sequence reads must therefore be joined together. \pb

The process of joining sequencing reads together is described as the
`assembly' stage in a genome project. Assembly software joins these numerous
sequence reads together into contiguous regions (`contigs') composed of many
overlapping individual reads. The goal of the assembly stage is produce
a continuous genome sequence describing the complete genome from start to
finish. \pb

Often repetitive nucleotide `repeat' regions in the genome or biased and
incomplete sequencing data may prevent the genome being assembled into
a continuous sequence. This sequence data can however describe the majority of
the genetic information in the genome and in the case of microbes and moving
forward with the current sequence data at this stage may be the best
investment of time and money \cite{branscomb2002}. \pb

A continuous high-quality `finished' genome sequence provides a much greater
depth of information as compared to a draft sequence
\cite{parkhill2002,fraser2002}. Examples of depth of information are complete
resolution of repeat regions and the exact distances between genomic elements.
The process of moving from a draft sequence to a finished continuous sequence
involves the process of `scaffolding' the contigs produced by the assembly
software together. \pb

>>>>

High throughput sequencing can produce hundreds of thousands to millions of
sequence reads from a genome. Each read represents the machine-interpreted
nucleotide sequence from a short region in the genome. At the time of writing
high throughput sequencing is limited to producing reads less than $>$1000
nucleotides. Therefore to produce to produce a sequence longer than this
length, e.g. a complete genome, these numerous individual reads must be pieced
together. The process of joining reads into longer sequences is the `assembly'
stage in a genome project \cite{assembly-reviews}. \pb

Assembly software takes the nucleotide reads produced by the sequencing
hardware and, in the ideal case, outputs a complete genome sequence compose of
these individual reads. An analogy to describe this is a jigsaw puzzle. Each
nucleotide read represents a single piece and the final genome sequence is the
completed puzzle. Repetitive nucleotide `repeat' regions within a genome or
biased and incomplete sequencing data may however prevent the genome being
assembled into a continuous sequence. This is analogous to missing jigsaw
pieces or pieces that fit to multiple other pieces. \pb

The recent advent of these current high throughput sequencing methods has lead
to a renewed interest into algorithms for solving the problem of genome
assembly \cite{assembly-reviews}. Genome assembly software is often unable to
completely these sequence reads together due to problems outlined above.
Instead assembly algorithms will generate large contiguous regions of sequence
(`contigs') composed of many individual reads. These represent optimum assembly
of reads into the longest sequences for the assembly algorithm. \pb

The process of ameliorating a draft sequence to a finished continuous sequence
can be expensive in time and laboratory effort. Therefore in many cases a set
of contigs may often describe enough of the genomic content for many research
questions \cite{branscomb2002}. A continuous high-quality `finished' genome
sequence does however provide further depth of information
\cite{parkhill2002,fraser2002}. The benefits of a complete genome sequence
include complete resolution of repeat regions and the exact distances between
genomic elements. This process of joining contigs from assembly together to form
a continuous sequence may be called `scaffolding' or `finishing'. Here on we
use the term scaffolding. \pb
